---
title: Isoplots and Trendplots
author: Tim Bergsma
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    keep_md : TRUE
    fig.width: 8
    fig.height: 6
    theme: united
    highlight: tango
vignette: >
  %\VignetteIndexEntry{Isoplots and Trendplots}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)
```

## Objective

In this exercise, we show some simple goodness-of-fit plots using isoplot()
and trendplot(). Admittedly, actual needs could quickly outstrip the flexibility
of these tools. Perhaps the concepts here can serve as a starting point
for further development.

## Data

We read example data from the yamlet package.

```{r, echo = TRUE}
library(qptools) 
library(magrittr)
library(yamlet)
library(ggplot2)
x <- 'extdata/modeling/xanomeline-mod.csv.gz' %>% system.file(package = 'yamlet') %>% io_csv %>% filter(cp > 1) %>% resolve
```

## DV vs IPRED by ACTARM per VISIT
 
```{r  DV-IPRED-ACTARM-VISIT, echo = TRUE, fig.width = 9.69, fig.height = 4.3, fig.cap = 'DV-IPRED-ACTARM-VISIT.png'}
#(
  x %>%
  ggplot(aes(IPRED, DV, color = ACTARM, pch = ACTARM)) + 
    geom_point(alpha = .5, size = 1.5) +
    scale_shape_manual(values = c(1,3))+
    facet_wrap(~VISIT, ncol = 3) +
  theme_bw() +
  theme(aspect.ratio = 1, legend.position = 'top', legend.title = element_blank()) +
  geom_abline(aes(slope = 1, intercept = 0))
  #) %>% devsize(3,3,verbose = T)
```

## Observations vs. Individual Predictions
 
```{r  DV-PRED, fig.width = 3.65, fig.height = 3.52, fig.cap = "x %>% isoplot(PRED, DV)"}
x %>% isoplot(PRED,DV)
```

## DV vs PRED log-log
```{r  DV-PRED-LOG, fig.width = 3.65, fig.height = 3.52, fig.cap = "x %>% isoplot(PRED, DV, transform = 'log10')"}
x %>% isoplot(PRED, DV, transform = 'log10')
```

## DV vs PRED and IPRED

```{r  DV-PRED-DV-IPRED, fig.width = 7.3, fig.height = 3.52, fig.cap = "x %>% isopair(PRED, DV, IPRED)"}
x %>% isopair(PRED, DV, IPRED)
```

## DV vs IPRED, untransformed and log-log
 
```{r  DV-IPRED-DV-IPRED-notrans-log, fig.width = 7.3, fig.height = 3.52, fig.cap = "x %>% isopair(IPRED, DV, transform = c('identity','log10'))"}
x %>% isopair(IPRED, DV, transform = c('identity','log10'))
```

## IWRES vs TIME
 
```{r  IWRES-TIME, fig.width = 3.65, fig.height = 3.52, fig.cap = "x %>% trendplot(TIME, IWRES)"}
x %>% trendplot(TIME, IWRES) #%>% devsize(3,3,verbose = T)
```

## IWRES vs TIME, untransformed and log-log
```{r  IWRES-TIME-IWRES-TIME-untrans-log, fig.width = 7.3, fig.height = 3.52, fig.cap = "x %>% trendpair(TIME, IWRES, transform_x = c('identity', 'log10'))"}
x %>% trendpair(TIME, IWRES, transform_x = c('identity', 'log10'))
```

